Visual Interrogation of Attention-Based Models: \\ Applications in Language Inference and Machine Comprehension

### 2018-05-28 EMNLP-demo-paper/NLPVis-demo-paper.tex
add figure
full intro/related work pass

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{emnlp2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}

@@

\begin{abstract}
Neural networks models have gained unprecedented popularity in natural language processing due to their state-of-the-art performance and the flexible end-to-end training scheme. Despite their advantages, the lack of interpretability hinders the further development and refinement of the model.
%These models often relying on the attention mechanism, which not only helps improve predicting performance but also provide interpretable representations for users to peak into the opaque models that are usually considered as black boxes.
%
In this work, we present a flexible visualization library for creating customized visual analytic environments, which enables the user to investigate and interrogate the relationship among the input, the model internals (i.e., attention), and the output predictions that in turn shed light on the model decision making processing.

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{emnlp2018}
\usepackage{times}
\usepackage{latexsym}
\usepackage{url}
\usepackage{mathptmx}

@@

\begin{abstract}
Neural networks models have gained unprecedented popularity in natural language processing due to their state-of-the-art performance and the flexible end-to-end training scheme. Despite their advantages, the lack of interpretability hinders the deployment and refinement of the model.
%These models often relying on the attention mechanism, which not only helps improve predicting performance but also provide interpretable representations for users to peak into the opaque models that are usually considered as black boxes.
%
In this work, we present a flexible visualization library for creating customized visual analytic environments, in which the user can investigate and interrogate the relationship among the input, the model internals (i.e., attention), and the output predictions that in turn shed light on the model decision making processing.

### 2018-05-28 EMNLP-demo-paper/intro.tex
full intro/related work pass

Deep neural networks have been successfully applied to various natural language processing tasks,
%where end-to-end training become a primary option.
and as the design of neural networks evolve, there is a clear trend of increasing model complexity both in term of architecture and parameters count. 
%
While the expressiveness of these models help improve the prediction performance, there is a fundamental lack of interpretability that lead many to consider neural network based models black boxes. 
%
Despite the difficulties, it is imperative to devise techniques to interpret the model internals and reason about a given prediction, which are essential to understand limitation of the model and improve upon the existing designs.

@@

Recently, attention networks have become a widely-adopted mechanism \cite{VaswaniShazeerParmar2017, Parikh2016}, the attention not only improves the model performance, but also yields interpretable intermediate representations (e.g. alignment among words). As a result, attention has been used as a natural interface for analyzing the internals of neural networks. Conduct analysis directly on the raw attention values can be challenging for human user, therefore, visual representations that highlight word alignments between sentences have been adopted, such as bipartite graph and heat map of the attention matrix~\cite{LiChenHovy2015, li2016understanding, lee2017interactive}.  However, many challenges remains. To begin with, for some NLP tasks, the word sequence pair among which the attentions are computed can be highly asymmetrical (i.e., in machine comprehension, the context paragraph can be much longer than the question sentence), where the standard visual encoding can not adequately handle.  In addition, many previous works present the attention in a statics setting. Yet, the ability to provide an interactive environment, where the user can instantaneously look at how changes in the input will affect the attention, and how small changes in attention will affect the prediction is crucial for interpreting the model. Finally, many previous visualization efforts, despite reveal many interesting results, focus more on illustrating potentially useful visualization technique than provide a flexible software tool that can be easily integrated into existing code base. As a result, applying these proposed techniques to real word examples may involving substantial engineering effort than many user may not willing to undertake.

To address these challenges and facilitate interpretation of attention-based models, 
we introduce a flexible python library that allow user to easily generate a web-based interactive visual analytics environment for various usage via a composable visualization component design.
%
To facilitate exploratory analysis (see Figure~\ref{fig:modelPipeline}), we adopted a fully interactive pipeline, where the user can perturb the input text (via two perturbation schemes discussed in Section~\ref{sec:perturb}) and observe the effects on attention and prediction, or modify the attention to understand how it affect the predictions.
%
We demonstrate our system on two primary NLP tasks: natural language
inference (NLI) and machine comprehension (MC). For the NLI task, we visualize
decomposable attention networks which represents a series of strongly performing models \cite{Parikh2016}.
For the MC task, we present your visualization methods on bidirectional attention flow model,
which also represents a line of strong models \cite{Seo2016}.

>>>>

Deep neural networks have been successfully applied to various natural language processing tasks,
%where end-to-end training become a primary option.
and as the design of neural networks evolve, there is a clear trend of increasing model complexity both in term of the architecture and the parameters count. 
%
While the expressiveness of these models help improve the prediction performance, there is a fundamental lack of interpretability that lead many to consider neural network models black boxes. 
%
Despite the difficulties, interpreting the model internals and reason about predictions are essential for understanding limitation of the model and improve upon the existing designs.

@@

Recently, attention networks have become a widely-adopted mechanism \cite{VaswaniShazeerParmar2017, Parikh2016}, the attention not only improves the model performance, but also yields interpretable intermediate representations (e.g. alignment among words). As a result, attention has been used as a natural interface for analyzing the internals of neural networks. Conduct analysis directly on the raw attention values can be challenging for human user, therefore, visual representations that highlight word alignments between sentences have been adopted, such as bipartite graph and heat map of the attention matrix~\cite{LiChenHovy2015, li2016understanding, lee2017interactive}.  However, many challenges remains. Firstly, for some NLP tasks, the word sequence pair among which the attentions are computed can be highly asymmetrical (i.e., in machine comprehension, the context paragraph can be much longer than the question sentence), where the standard visual encoding can not adequately handle.  In addition, many previous works present the attention in a statics setting. Yet, the ability to provide an interactive environment, where the user can instantaneously look at how changes in the input will affect the attention, and how small changes in attention will affect the prediction are crucial for interpreting the model. Finally, many previous visualization efforts, despite reveal many interesting results, focus more on illustrating potentially useful visualization technique than provide a flexible software tool that can be easily integrated into existing code base. As a result, applying these proposed techniques to real word examples may involving substantial engineering effort than many user may not affort to undertake.

To address these challenges in interpretating attention-based models, 
we introduce a flexible python library that allow user to easily generate a web-based interactive visual analytics environment for various usage via a composable visualization component design.
%
To facilitate exploratory analysis (see Figure~\ref{fig:modelPipeline}), we adopted a fully interactive pipeline, where the user can perturb the input text (via two perturbation schemes discussed in Section~\ref{sec:perturb}) and observe the effects on attention and prediction, or modify the attention to understand how it affect the predictions.
%
We demonstrate our system on two primary NLP tasks: natural language
inference (NLI) and machine comprehension (MC). For the NLI task, we visualize
decomposable attention networks which represents a series of strongly performing models \cite{Parikh2016}.
For the MC task, we present your visualization methods on bidirectional attention flow model,
which also represents a line of state-of-the-art models \cite{Seo2016}.

### 2018-05-28 EMNLP-demo-paper/relatedWorks.tex
full intro/related work pass

\section{Related Works}

Visualizing and understanding neural models in nlp \cite{LiChenHovy2015}
Understanding Neural Networks through Representation Erasure \cite{li2016understanding}

AllenNLP demo

%Interactive Visualization and Manipulation of Attention-based Neural Machine Translation
Lee et al. visualizes the beam search and attention component in the NMT model~\cite{lee2017interactive}. The work allows user involve into the decision-making process by enabling dynamic change the probability for the next step of the search tree or change the weight value of the attention.
%}

%End-to-End Non-Factoid Question Answering with an Interactive Visualization of Neural Attention Weights
%\textcolor{blue}{
In the recent attention visualization work focus on question and answering model~\cite{ruckle2017end}, 
instead of visualizing the attention matrix directly, the system shows the text context and highlights the critical phrase, which used to answers the questions, using color map. In this system, the user is able to use the visual feature to compare the benefit and drawback of multiple attention based neural network model.
%}

>>>>

\section{Related Works}
Due to the increasing demand for model interpretability, many previous works has been proposed for understanding NLP models in term of making sense of individual predictions as well as model mechanism as a whole.

In the recent work~\cite{LiChenHovy2015}, the composability of the vector-based representation of text is investigate by utilizing instance based attribution techniques originated from the vision community (i.e., \cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017}). 
%
In the representation erasure work~\cite{li2016understanding}, the authors explain the neural model decisions by exploring the impact of various parts of the model by altering or removing the corresponding representations (i.e., the dimension count of hidden units, or input words).

Beside interpreting the model via carefully designed experiments, interactive demo / visualization systems, such as AllenNLP demo (\url{http://demo.allennlp.org/}), also attract attention as it provide a flexible environment where the user can experiment and explore the various input combination and carry out error analysis.
%
Some of these interactive systems incorporate various visual encodings to summary the key information of the model.
%
The training dynamic of RNN network is visualized in the RNNbow system~\cite{CashmanPattersonMosca2017}, and the hidden state properties of the LSTM are investigated in the LSTMvis visualization tool~\cite{StrobeltGehrmannPfister2018}.
%Interactive Visualization and Manipulation of Attention-based Neural Machine Translation
Lee et al. visualizes the beam search and attention component in the NMT model. The work~\cite{lee2017interactive} directly involve the user into the decision-making process by enabling dynamic change the probability for the next step of the search tree or change the weight value of the attention.
%}
In the recent attention visualization work on question and answering model~\cite{ruckle2017end}, 
instead of visualizing the attention matrix directly, the system using color map to show the text context and highlights the critical phrase that used to answers the questions. %The user is able to use the visual feature to compare the benefit and drawback of multiple attention based neural network model.
%
